Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-03-27 document/ms.tex
catalog notation

% Document specific variables.
\newcommand{\rate}{\ensuremath{\eta}}
\newcommand{\ratepars}{\ensuremath{\theta}}
\newcommand{\obs}[1]{\ensuremath{\hat{#1}}}
\newcommand{\radius}{\ensuremath{R}}
\newcommand{\period}{\ensuremath{T}}
\newcommand{\completeness}{\ensuremath{P_\mathrm{c}}}
\newcommand{\transitprob}{\ensuremath{P_\mathrm{t}}}

\newcommand{\interim}{\ensuremath{\alpha}}

@@

The purpose of this \paper\ is to determine constraints on the \emph{rate
function} of exoplanets
\begin{eqnarray}\eqlabel{true-rate}
\rate_\ratepars (\period,\,\radius)
&=& \frac{\dd N}{\dd\ln\period\dd\ln\radius}
\end{eqnarray}
given our dataset.
Here we have not made any assumptions about the functional form of \rate, we
are only stating that it can be modeled by some---possibly very large---set of
parameters \ratepars.

In this \paper, we will assume that the completeness function of the survey is
known and it can be represented as a probability at every (true) period and
radius $\completeness (\period,\,\radius)$.
Similarly, we can consider the marginalized transit probability $\transitprob
(\period)$ as only a function of period.
These selection effects can be incorporated into the rate from \eq{true-rate}
to get the \emph{observed rate}
\begin{eqnarray}\eqlabel{obs-rate}
\obs{\rate} (\period,\,\radius;\,\ratepars)
&=& \rate (\period,\,\radius;\,\ratepars)\,\completeness(\period,\,\radius)\,
\transitprob(\period)\quad.
\end{eqnarray}

Now we will make the further assumption that every candidate in our sample of
detected exoplanets is an independent Poisson sample from the population
described by our rate function.
Under this assumption, the likelihood of the catalog conditioned on a set of
parameters \ratepars\ is (see Appendix XXXXX for a derivation)
\begin{eqnarray}
\ln p(\period,\,\radius\,|\,\ratepars) &\propto&
\sum_{k=1}^K \ln\obs{\rate} (\period_k,\,\radius_k;\,\ratepars)
- \int \obs{\rate} (\period,\,\radius;\,\ratepars) \dd\period\dd\radius
\quad.
\end{eqnarray}

@@

>>>>

% Document specific variables.
\newcommand{\rate}{\ensuremath{\eta}}
\newcommand{\ratepars}{{\ensuremath{\bvec{\theta}}}}
\newcommand{\obs}[1]{\ensuremath{\hat{#1}}}
\newcommand{\radius}{\ensuremath{R}}
\newcommand{\period}{\ensuremath{T}}
\newcommand{\completeness}{{\ensuremath{P_\mathrm{c}}}}
\newcommand{\transitprob}{{\ensuremath{P_\mathrm{t}}}}
\newcommand{\entry}{\ensuremath{S}}
\newcommand{\catalog}{{\ensuremath{\bvec{\entry}}}}

\newcommand{\interim}{{\ensuremath{\bvec{\alpha}}}}

@@

Let's also introduce more succinct notation for a \emph{catalog}.
We'll refer to the set of period and radius measurements for a set of $K$
candidate exoplanets as
\begin{eqnarray}
\catalog &=& \{\entry_k\}_{k=1}^K
\end{eqnarray}
where each entry $\entry_k = (\period_k,\,\radius_k)$ is a tuple containing a
period and radius.

The purpose of this \paper\ is to determine constraints on the \emph{rate
function} of exoplanets
\begin{eqnarray}\eqlabel{true-rate}
\rate_\ratepars (\entry)
&=& \frac{\dd N}{\dd\ln\period\dd\ln\radius}
\end{eqnarray}
given our dataset.
Here we have not made any assumptions about the functional form of \rate, we
are only stating that it can be modeled by some---possibly very large---set of
parameters \ratepars.

In this \paper, we will assume that the completeness function of the survey is
known and it can be represented as a probability at every (true) period and
radius $\completeness (\entry)$.
Similarly, we can consider the marginalized transit probability $\transitprob
(\period)$ as only a function of period.
These selection effects can be incorporated into the rate from \eq{true-rate}
to get the \emph{observable rate}
\begin{eqnarray}\eqlabel{obs-rate}
\obs{\rate}_\ratepars (\entry)
&=& \rate_\ratepars(\entry)\,\completeness(\entry)\,
\transitprob(\period)\quad.
\end{eqnarray}

Now we will make the further assumption that every candidate in our sample of
detected exoplanets is an independent Poisson sample from the population
described by our rate function.
Under this assumption, the likelihood of a catalog \catalog\ conditioned on a
set of parameters \ratepars\ is (see Appendix XXXXX for a derivation)
\begin{eqnarray}\eqlabel{true-like}
\ln p(\catalog\,|\,\ratepars) &\propto&
\sum_{k=1}^K \ln\obs{\rate}_\ratepars (\entry_k)
- \int \obs{\rate}_\ratepars (\entry) \dd\entry
\quad.
\end{eqnarray}

@@

The ``likelihood function'' given in \eq{true-like}\ldots
